---
title: "Coursera Statistical Inference Peer Assesment Part 2"
output: 
  html_document:
    keep_md: true
---

# Coursera Statistical Inference Peer Assesment - Part 2

<!-- Assignment 

1. Load the ToothGrowth data and perform some basic exploratory data
analyses

2. Provide a basic summary of the data.  

3. Use confidence intervals and hypothesis tests to compare tooth growth
by supp and dose. (Use the techniques from class even if there's other
approaches worth considering)

4. State your conclusions and the assumptions needed for your
conclusions.  -->

## ToothGrowth Data - BASIC EXPLORATION

### Read in the data and plot
```{r}
data("ToothGrowth")
par(mfrow=c(1,2)) 
for (supp in c("OJ","VC")) {
    plot(ToothGrowth[ToothGrowth$supp == supp,]$dose, 
    		ToothGrowth[ToothGrowth$supp == supp,]$len, 
		xlab="dose", ylab="length", main=supp)
}
```

We can see that despite sigificant variation in each group, there does appear to be a dependence of the tooth size on the dosage. Whether a dependence on the method of delivery (OJ vs VC) exists is less clear.

### Basic Summary of the Data

#### Sample Mean and Variance
```{r}
mean(ToothGrowth$len)
sd(ToothGrowth$len)
```

#### Sample Means and standard deviations for each dose and delivery method
```{r}
for (supp in c("OJ","VC")) {
    for (dose in c(0.5,1.0,2.0)) {
    	v <-  ToothGrowth[ToothGrowth$supp == supp & ToothGrowth$dose == dose,]$len
	print (mean (v))
	print (sd(v))
    }
}

```

To visualize these statistics, boxplot is useful:

```{r}
par(mfrow=c(1,2)) 
boxplot(len ~ supp,data=ToothGrowth, main="Length",
   xlab="supp", ylab="length") 
boxplot(len ~ dose,data=ToothGrowth, main="Length",
   xlab="dose", ylab="length") 
```


### Compare Tooth length by supp and dose
This is perhaps more properly modelled by a regression analysis. We will see if increasing the dosage will have an effect on the tooth lenghts. For this, we'll take vectors of differences in the tooth lenghts for dose = 0.5 and dose = 1.0, as well as dose = 1.0 and dose = 2.0.
Our null hypothesis in both cases are that the mean of the differences is zero, i.e. increasing the dosage has no effect.
```{r}
d_oj_12 <- ToothGrowth[ToothGrowth$supp == 'OJ' & 
	ToothGrowth$dose == 1.0,]$len - 
	ToothGrowth[ToothGrowth$supp == 'OJ' & 
	ToothGrowth$dose == 0.5,]$len 
	
d_oj_23 <- ToothGrowth[ToothGrowth$supp == 'OJ' & 
	ToothGrowth$dose == 2.0,]$len - 
	ToothGrowth[ToothGrowth$supp == 'OJ' & 
	ToothGrowth$dose == 1.0,]$len 
	
d_vc_12 <- ToothGrowth[ToothGrowth$supp == 'VC' & 
	ToothGrowth$dose == 1.0,]$len - 
	ToothGrowth[ToothGrowth$supp == 'VC' & 
	ToothGrowth$dose == 0.5,]$len 

d_vc_23 <- ToothGrowth[ToothGrowth$supp == 'VC' & 
	ToothGrowth$dose == 2.0,]$len - 
	ToothGrowth[ToothGrowth$supp == 'VC' & 
	ToothGrowth$dose == 1.0,]$len 

d_oj_13 <- ToothGrowth[ToothGrowth$supp == 'OJ' & 
	ToothGrowth$dose == 2.0,]$len - 
	ToothGrowth[ToothGrowth$supp == 'OJ' & 
	ToothGrowth$dose == 0.5,]$len 
	
d_vc_13 <- ToothGrowth[ToothGrowth$supp == 'VC' & 
	ToothGrowth$dose == 2.0,]$len - 
	ToothGrowth[ToothGrowth$supp == 'VC' & 
	ToothGrowth$dose == 0.5,]$len 
```

I will only include R output for two t tests in this report, the rest can be reproduced trivially.

```{r}
t.test(d_vc_12)
```
The 95% confidence interval does not contain zero, we therefore reject the null null hypothesis and conclude that the tooth length does appear to depend on the dose. 

However, as can be seen from the differences between dose=2.0 and dose=1.0 for the OJ delivery, the effects of increasing the dosage may diminish rapidly. In that case, we failed to reject the null hypothesis.
```{r}
t.test(d_oj_23)
```

We can also test if the two methods of delivery produce statistically significant results. We'll consider the vector of differences between d_oj_13 and d_vc_13, with null hypothesis that the mean is zero.

```{r}
t.test(d_vc_13-d_oj_13)
```

Again, we fail to reject the null hypothesis confidence 95%, since the corresponding t interval contains zero. The evidence does not suport the claim that VC is more effective in promoting tooth growth.

### Conclusions and Assumptions
We assumed that the distribution of the tooth lengths has central symmetry, in order to use Student's t test. 