---
title: "Coursera Statistical Inference Peer Assesment"
output: 
  html_document:
    keep_md: true
---

# Coursera Statistical Inference Peer Assesment

<!-- Assignment
# The exponential distribution can be simulated in R with rexp(n,
# lambda) where lambda is the rate parameter. The mean of exponential
# distribution is 1/lambda and the standard deviation is also also
# 1/lambda. Set lambda = 0.2 for all of the simulations. In this
# simulation, you will investigate the distribution of averages of 40
# exponential(0.2)s. Note that you will need to do a thousand or so
# simulated averages of 40 exponentials.

Illustrate via simulation and associated explanatory text the
properties of the distribution of the mean of 40 exponential(0.2)s.
You should

1. Show where the distribution is centered at and compare it to the
theoretical center of the distribution.
2. Show how variable it is and compare it to the theoretical variance
of the distribution.
3. Show that the distribution is approximately normal.
4. Evaluate the coverage of the confidence interval for 1/lambda: \bar
X ±1.96Sn.

-->

## Experiments with the exponential distibution

### Set up the experiment run
```{r}
lambda <- 0.2
n <- 40
N <- 2^16 # approx. 65k experiments
```
#### Run N experiments, each consisting of n draws from the $exp(\lambda)$ distribution.
```{r}
f <- function (x) rexp(n,lambda)
E <- sapply(1:N,f)
sample_means <- (apply(E,2,mean))
length(sample_means)
```
### Visualize Results
```{r}
hist(sample_means,freq=FALSE,prob=TRUE,main="distribution of sample means")
lines(density(sample_means),col='blue')

# Plot a normal distribution with the theoretically-predicted parameters
x <- seq(min(sample_means), max(sample_means), length=N)
hx <- dnorm(x,mean=1/lambda,sd=1/(lambda*sqrt(n)))
lines(x, hx, type="l", lty=2,col='red')

# legend
legend(6.,0.45,c("observed sample means","normal approximation"),lwd=c(2.5,2.5),col=c("blue","red"))
```

### Distribution center
The expected value of the sample means is the same as the mean of the underlying random variable:
$$ \mathbf{E} (\bar x) = \bar X = \frac{1}{\lambda} = \frac{1}{0.2} = 5.0$$
```{r}
mean(sample_means)
```

### Observed Variance of the Samples
The variance of the sample means is $$Var(\bar X) = \frac{Var(X)}{n} = \frac{1/\lambda^2}{n} = 25/40 = 0.625$$ The standard deviation is correspondingly $$\sqrt{Var(\bar X)} = \frac{1}{\lambda\sqrt{n}} = \frac{5}{\sqrt{40}} \approx 0.7905694$$

```{r}
sd(sample_means)^2 # variance
sd(sample_means) # standard deviation
```

### CLT: Limit is Normal
The 'standard' way of comparing the sample distribution with the Normal is a Q-Q plot, which gives a comparison of the cumulative distribution function to that of the Gaussian.
```{r}
qqnorm(sample_means)
qqline(sample_means)
```

We see a good match in the region of about $\sigma$ around the mean, and a poorer match on the fringes of the distribution. This is not surprising, considering that the exponential distribution(and hence our samples), unlike the normal, only attains positive values. A better illustration of the CLT is given by drastically larger samples, e.g. for n~1k:
```{r}
g <- function(x) rexp(2^10,lambda)
sample_means_2 <- (apply(sapply(1:N,g),2,mean))
qqnorm(sample_means_2)
qqline(sample_means_2)
```

### Coverage of the confidence interval for $1/\lambda \in \bar X \pm 1.96 \frac{s}{\sqrt{n}}$
We are going to count the number of our sample means that fall inside 1.96 estimated sample standard deviations from the estimated sample mean:
```{r}
hits <- sum(rep(1,length(sample_means[
     sample_means >= mean(sample_means)-1.96*sd(sample_means) 
     & sample_means <= mean(sample_means)+1.96*sd(sample_means)])))
hits
```
Which is 95% of the total number of trials we performed:
```{r}
hits/N
```

# The End