---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
x <- read.csv("./activity.csv")
x$date <- as.Date.character(x$date)
```

## What is mean total number of steps taken per day?

### Histogram of the total number of steps taken each day
```{r}
dates <- unique(x$date)
stepsDay <- numeric()
for (d in dates) {
  stepsDay <- c(stepsDay,sum(x[x$date == d,]$steps) )
}
plot(dates,stepsDay,"h")
```

### mean and median total number of steps taken per day
```{r}
mean(stepsDay,na.rm=TRUE)
median(stepsDay,na.rm=TRUE)
```


## What is the average daily activity pattern?

###  the average number of steps taken, averaged across all days, vs 5-minute interval

```{r}
intervals <- unique(x$interval)
stepsInterval <- numeric()
for (i in intervals) {
  stepsInterval <- c(stepsInterval, mean(x[x$interval==i,]$steps,na.rm=TRUE))
}
plot(intervals,stepsInterval,"l")
```

### Find the 5-minute interval with maximum average number of steps

```{r}
intervals[which.max(stepsInterval)]
```

## Imputing missing values

### total number of missing values in the dataset
```{r}
nrow(x[is.na(x$steps),])
```

### Fill in the missing data in the dataset

some prep work
```{r}
y <-x
si <- cbind.data.frame(intervals,stepsInterval)
```

We will replace the missing values by the average value for the relevant time interval

```{r}
for (r in as.numeric(rownames(y))) {
  if (is.na(y[r,"steps"])) {
    y[r,"steps"] <- si[si$intervals==y[r,"interval"],"stepsInterval"]
  }
}
```

### New histogram of the daily steps taken

```{r}
stepsDayNew <- numeric()
for (d in dates) {
  stepsDayNew <- c(stepsDayNew,sum(y[y$date == d,]$steps) )
}
plot(dates,stepsDayNew,"h")
```

### New mean and median

```{r}
mean(stepsDayNew,na.rm=FALSE)
median(stepsDayNew,na.rm=FALSE)
```

## Are there differences in activity patterns between weekdays and weekends?

### augment the filled-in data set with weekend/weekday indicator

```{r}
isweekend <- weekdays(y$date) %in% c("Saturday","Sunday")
y <- cbind(y,isweekend)
```

### plot the 5-minute interval averages for weekdays and weekends

```{r}
stepsIntervalWeekend <- numeric()
stepsIntervalWeekday <- numeric()
for (i in intervals) {
  stepsIntervalWeekend <- c(stepsIntervalWeekend, mean(y[y$isweekend & y$interval==i,]$steps,na.rm=TRUE))
  stepsIntervalWeekday <- c(stepsIntervalWeekday, mean(y[(! y$isweekend ) & y$interval==i,]$steps,na.rm=TRUE))
}
par(mfrow=c(2,1))
plot(intervals,stepsIntervalWeekend,"l",main="Weekends")
plot(intervals,stepsIntervalWeekday,"l",main="Weekdays")
```

## The End